\documentclass[12pt]{article}
\input{nets.sty}

\title{Nets}
\author{Austin Gill}

\begin{document}
\maketitle

We have already seen that sequences are ``adequate'' to detect limit points, continuous functions,
and compact sets in metrizable spaces. There is a generalization of the notion of a sequence,
called a \textit{net}, that will do the same for an arbitrary topological space.

\section{Directed Sets}
\begin{defn}
    A relation $\preceq$ on a set $A$ is called a \textbf{partial order} relation if the following
    conditions hold:
    \begin{enumerate}
        \item $\alpha \preceq \alpha$ for all $\alpha \in A$.
        \item If $\alpha \preceq \beta$ and $\beta \preceq \alpha$, then $\alpha = \beta$.
        \item If $\alpha \preceq \beta$ and $\beta \preceq \gamma$, then $\alpha \preceq \gamma$.
    \end{enumerate}
\end{defn}

\begin{defn}
    A \textbf{directed set} $J$ is a set with a partial order $\preceq$ such that for each pair
    $\alpha, \beta$ of elements of $J$, there exists an element $\gamma \in J$ with the property
    that
    $\alpha \preceq \gamma$ and $\beta \preceq \gamma$.
\end{defn}
Show the following are directed sets.

\begin{enumerate}
    \item Any simply ordered set under the relation $\leq$.
          \begin{proof}
              Let $A$ be a simply ordered set with order relation $<$. Show that the set $A$ is
              directed under the relation $\leq$.

              We need to show that
              \begin{enumerate}
                  \item\label{pf:part-order} $A$ is partially ordered by $\leq$.

                        Let $a, b, c \in A$.
                        \begin{enumerate}
                            \item $a \leq a$ is trivially true, because $a = a$!
                            \item Let $a \leq b$ and $b \leq a$. Because $A$ is simply ordered, $a
                                      < b$ and $b < a$ cannot hold, so the only other possibility
                                  is that $a = b$.
                            \item Let $a \leq b$ and $b \leq c$. Then clearly $a \leq c$.
                        \end{enumerate}
                        So therefore $A$ is partially ordered by $\leq$.
                  \item\label{pf:part-directed} For every $a, b \in A$ there exists a $c \in A$
                        such that $a \leq c$ and $b \leq c$.

                        Let $a, b \in A$. We wish to show that there exists a $c \in A$ with $a
                            \leq c$ and $b \leq c$.

                        $A$ is simply ordered, so either $a < b$ or $b < a$. Without loss of
                        generality, let $a < b$. But from (\ref{pf:part-order}) $A$ is partially
                        ordered by $\leq$, so we
                        have $b \leq b$!
              \end{enumerate}
              Then $\leq$ is a partial ordering on $A$ and for every pair $a, b \in A$ we have
              either $a \leq b$ and $b \leq b$, or we have $b \leq a$ and $a \leq a$! Thus $A$ is
              directed.
          \end{proof}

    \item The collection of all subsets of a set $S$, partially ordered by inclusion. That is, $A
              \preceq B$ if $A \subset B$.
          \begin{proof}
              Let $A, B \in \mathcal P(S)$. We wish to show that $\mathcal P(S)$ is directed, that
              is, that there exists a $C \in \mathcal P(S)$ with $A \subset C$ and $B \subset C$.

              Note that $S \in \mathcal P(S)$ and that $A \subset S$ and $B \subset S$. Thus
              $\mathcal P(S)$ is a directed set under inclusion.
          \end{proof}

    \item\label{prob:1c} The collection $\mathcal A$ of subsets of $S$ that is closed under finite
          intersections,
          partially ordered by reverse inclusion. That is, $A \preceq B$ if $A \supset B$.
          \begin{proof}
              Let $\mathcal A \subset \mathcal P(S)$ such that if $A_1, \dots, A_n \in \mathcal A$
              then $\bigcap_{i=1}^n \in \mathcal A$, and let $A_1, A_2 \in \mathcal A$. We wish to
              show that $\mathcal A$ is directed. That is, that there exists a $B \in \mathcal A$
              with $A_1 \supset B$ and $A_2 \supset B$.

              But we know $\mathcal A$ is closed under finite intersections, so
              $A_1 \cap A_2 \in \mathcal A$. Moreover, we know that $A_1 \supset A_1 \cap A_2$ and
              $A_2 \supset A_1 \cap A_2$. Thus $\mathcal A$ is directed.
          \end{proof}

    \item The collection of all closed subsets of a space $X$, partially ordered by inclusion.
          \begin{proof}
              Let $\mathcal A$ be the collection of all closed subsets of a space $X$.
              Let $A, B \in \mathcal A$, that is, let $A$ and $B$ be closed in $X$.
              We wish to show that $\mathcal A$ is directed. That is, that there exists a $C \in
                  \mathcal A$ with $A \subset C$ and $B \subset C$.

              Recall that finite unions of closed sets are closed, so $A \cup B$ is closed.
              Moreover, note that $A \subset A \cup B$ and $B \subset A \cup B$.
              Thus $\mathcal A$ is directed.
          \end{proof}
\end{enumerate}

\section{Cofinal Sets}
\begin{defn}
    A subset $K \subset J$ is said to be \textbf{cofinal} in $J$ if for each $\alpha \in J$, there
    exists a $\beta \in K$ such that $\alpha \preceq \beta$.
\end{defn}

Show that if $J$ is directed, and $K$ is cofinal in $J$, then $K$ is directed.

\begin{proof}
    Let $J$ be directed and $K$ be a cofinal subset of $J$. Then for each $a, b \in J$
    there exists a $c \in J$ with $a \preceq c$ and $b \preceq c$, and for each
    $a \in J$ there exists a $\beta \in K$ such that $a \preceq \beta$. We wish
    to show that for each $\alpha, \beta \in K$ there exists a $\gamma \in K$ with
    $\alpha \preceq \gamma$ and $\beta \preceq \gamma$.

    So let $\alpha, \beta \in K$. Since $K \subset J$, $\alpha, \beta \in J$. Since $J$ is
    directed, there is a $c \in J$ with $\alpha, \beta \preceq c$. But $c \in J$ and $K$ is cofinal
    in
    $J$, so there is a $\gamma \in K$ with $c \preceq \gamma$. Since $\alpha, \beta \preceq c$ and
    $c \preceq \gamma$, we have that $\alpha, \beta \preceq \gamma$. That is, that $K$ is directed.
\end{proof}

\section{Nets and Convergence}
\begin{defn}
    Let $X$ be a topological space. A \textbf{net} in $X$ is a function $f$ from a directed set $J$
    into $X$. If $\alpha \in J$, we usually denote $f(\alpha)$ by $x_\alpha$. We denote the net $f$
    itself by ${\left(x_\alpha\right)}_{\alpha \in J}$, or simply $(x_\alpha)$ if $J$ is
    understood.
\end{defn}
\begin{defn}
    The net $(x_\alpha)$ is said to \textbf{converge} to a point $x \in X$, denoted by $x_\alpha
        \to x$ if for each neighborhood $U$ of $x$, there exists an $\alpha \in J$ such that

    \[\alpha \preceq \beta \Longrightarrow x_\beta \in U\]
\end{defn}

Show that these definitions are equivalent to those of a sequence and the convergence of a sequence
when $J = \Z_+$.

\begin{defn}
    Given a set $X$, we define a \textbf{sequence} of elements of $X$ to be a function $\vec x :
        \Z_+ \to X$.
\end{defn}

\begin{defn}
    A sequence $(x_n)$ \textbf{converges} to a point $x \in X$ if for every neighborhood $U$ of
    $x$, there is an $n^* \in \mathbb \Z_+$ such that $x_n \in U$ for all $n \geq n^*$.
\end{defn}

\begin{proof}
    Let $X$ be a topological space and $f : \Z_+ \to X$ be a net in $X$. Then clearly $f$ is a
    sequence.

    Let $X$ be a topological space and $\vec x : \Z_+ \to X$ be a sequence of elements of $X$. Then
    clearly $\vec x$ is also a net.
\end{proof}
\begin{proof}
    Let $X$ be a topological space and let $f : \Z_+ \to X$ be a net that converges to $x \in X$.
    That is, for every neighborhood $U$ of $x$, there exists an $n^* \in \Z_+$ such that
    \[ n^* \preceq n \Longrightarrow x_n \in U \]
    But the partial ordering $\preceq$ of $\Z_+$ is $\leq$, so this is precisely the definition of
    sequence convergence.
\end{proof}

\section{Convergence in a Product Space}
Suppose that $(x_\alpha)$ converges to $x$ in $X$ and $(y_\alpha)$ converges to $y$ in $Y$. Show
that $(x_\alpha \times y_\alpha)$ converges to $x \times y$ in $X \times Y$.
\begin{proof}
    Let $(x_\alpha) \to x$ in $X$ and $(y_\alpha) \to y$ in $Y$, and let $U \times V$ be a
    neighborhood of $x \times y$ in $X \times Y$.

    Since $(x_\alpha) \to x$ in $X$, there is an $\alpha \in J$, the index set, with $x_i \in U$
    for all $\alpha \preceq i$. Likewise, since $(y_\alpha) \to y$ in $Y$, there is a $\beta \in J$
    with $x_i \in V$ for all $\beta \preceq i$.

    But $J$ is directed, so there is a $\gamma \in J$ with $\gamma \preceq i$ implying both
    $\alpha \preceq i$ and $\beta \preceq i$. Then for all $\gamma \preceq i$, we have that
    $x_i \times y_i \in U \times V$. Therefore $(x_\alpha \times y_\alpha) \to x \times y$
    in $X \times Y$.
\end{proof}

\section{Convergence in a Hausdorff Space}
Show that if $X$ is Hausdorff, a net in $X$ converges to at most one point.
\begin{proof}
    Let $X$ be a Hausdorff topological space, and let $(x_\alpha)$ be a net in $X$.

    Suppose by way of contradiction that $(x_\alpha) \to x_0$ and $(x_\alpha) \to y_0$ with
    $x_0 \neq y_0$. Since $X$ is Hausdorff, let $U$ and $V$ be disjoint neighborhoods of
    $x_0$ and $y_0$ respectively. Then there exist some $\alpha, \beta \in J$ with
    $x_i \in U$ and $x_j \in V$ for all $\alpha \preceq i$ and $\beta \preceq j$.

    Since $J$ is directed, there is a $\gamma \in J$ for which $\gamma \preceq i$ implies both
    $\alpha \preceq i$ and $\beta \preceq i$, so then $x_i \in U$ and $x_i \in V$. But this
    contradicts the fact that $U$ and $V$ are disjoint.

    Therefore $(x_\alpha)$ converges to at most one point of $X$.
\end{proof}

\section{Closure of a Set}
\begin{thm}
    Let $A \subset X$. Then $x \in \overline A$ if and only if there is a net of points of $A$
    converging to $x$.
\end{thm}
\begin{proof}
    % Move the \item to the next line...
    \ \vspace{-0.5cm}\newline
    \begin{itemize}
        \item[$\lbrack\Rightarrow\rbrack$] Let $A \subset X$ and suppose $x \in \overline A$. We
              wish to show that there is a net in $A$ that converges to $x$.

              For the sheer hell of it, let $J$ be the collection of all neighborhoods of $x$,
              partially ordered by reverse inclusion. Since $J$ is a collection of open
              neighborhoods, it is closed under finite intersections. Then by problem 1.3 above,
              $J$ is directed.

              Since $x \in \overline A$, every $U \in J$ intersects $A$. So choose a point
              $x_U \in U \cap A$ for every $U \in J$. Then we claim that $(x_U)$ is a net in
              $A$ that converges to $x$.

              So let\footnote{I believe\texttrademark{} that $U$ plays the role of
                  both $U$ and $\alpha$ from the definition of net convergence. We can do this
                  because it's both a neighborhood and an element of the index set.}
              $U \in J$. Then for any $V \subset U$, or rather $U \supset V$, that is
              $U \preceq V$, we have a $x_V \in V \subset U$. That is,
              $U \preceq V \Rightarrow x_V \in U$. Thus $(x_U) \to x$.
        \item[$\lbrack\Leftarrow\rbrack$] Let $(x_\alpha) \to x$ be a net in $A$ converging to $x$.
              We wish to show that $x \in \overline A$.

              So then for every neighborhood $U$ of $x$, there exists an $\alpha \in J$ such that
              if $\alpha \preceq \beta$, then $x_\beta \in U$. But $(x_\alpha)$ is a net of points
              of $A$ and $\alpha \preceq \alpha$ for every $\alpha$, so $x_\alpha$ is in $A$ and
              thus $x_\alpha \in U$. That is, for every neighborhood $U$ of $x$,
              $U \cap A \neq \varnothing$. Thus $x \in \overline A$.
    \end{itemize}
\end{proof}

\section{Continuity}
\begin{defn}
    A function $f : X \to Y$ is said to be \textbf{continuous} if for each open subset $V$ of $Y$,
    the set $f^{-1}(V)$ is an open subset of $X$.
\end{defn}
\begin{thm}
    Let $f: X \to Y$. Then $f$ is continuous if and only if for every convergent net $(x_\alpha)$
    in $X$ converging to a point $x$, the net $\big(f(x_\alpha)\big)$ converges to $f(x)$.
\end{thm}
\begin{proof}
    % Move the \item to the next line...
    \ \vspace{-0.5cm}\newline
    \begin{itemize}
        \item[$\lbrack\Rightarrow\rbrack$] Let $f: X \to Y$ be a continuous function and suppose
              $(x_\alpha) \to x$. We wish to show the net $\big(f(x_\alpha)\big) \to f(x)$, that
              is, for every neighborhood $V$ of $f(x)$, there exists an $\alpha \in J$ such that
              if $\alpha \preceq \beta$, then $f(x_\beta) \in V$.

              So let $V \subset Y$ be a neighborhood of $f(x)$. Since $f$ is continuous, then the
              preimage $f^{-1}(V)$ is open in $X$. But $(x_\alpha) \to x$, so for
              neighborhood $U$ of $x$, there exists an $\alpha \in J$ such that if
              $\alpha \preceq \beta$, then $x_\beta \in U$.

              So then there exists an $\alpha \in J$ such that if $\alpha \preceq \beta$, then
              $x_\beta \in f^{-1}(V)$ but if $x_\beta \in f^{-1}(V)$, then $f(x_\beta) \in V$.
              Therefore $\big(f(x_\alpha)\big) \to f(x)$.
        \item[$\lbrack\Leftarrow\rbrack$]
    \end{itemize}
\end{proof}

\section{Subnets}
\begin{defn}
    Let $f: J \to X$ be a net in $X$; let $f(\alpha) = x_\alpha$. If $K$ is a directed set and $g:
        K \to J$ is a function such that
    \begin{enumerate}
        \item $i \preceq j \Rightarrow g(i) \preceq g(j)$
        \item $g(K)$ is cofinal in $J$
    \end{enumerate}
    then the composite function $f \circ g : K \to X$ is called a \textbf{subnet} if $(x_\alpha)$.
\end{defn}

Show that if the net $(x_\alpha)$ converges to $x$, so does any subnet of $(x_\alpha)$.
\begin{proof}
    \todo{Prove.}
\end{proof}

\section{Accumulation Points}
\begin{defn}
    Let ${(x_\alpha)}_{\alpha\in J}$ be a a net in $X$. We say that $x$ is an \textbf{accumulation
        point} of the net if for each neighborhood $U$ of $x$, the set of $\alpha$ for which
    $x_\alpha \in U$ is cofinal in $J$.
\end{defn}

\begin{lemma}
    The net $(x_\alpha)$ has the point $x$ as an accumulation point if and only if some subnet of
    $(x_\alpha)$ converges to $x$.
\end{lemma}
\begin{proof}
    \todo{To prove $\Rightarrow$, let $K$ be the set of all pairs $(\alpha, U)$ where $\alpha \in
            J$ and $U$ is a neighborhood of $x$ containing $x_\alpha$. Define $(\alpha, U) \preceq
            (\beta, V)$
        if $\alpha \preceq \beta$ and $V \subset U$. Show that $K$ is directed and use it to define
        the subnet.}
\end{proof}

\section{Compactness}
\begin{thm}
    $X$ is compact if and only if every net in $X$ has a convergent subnet.
\end{thm}
\begin{proof}
    \todo{To prove $\Rightarrow$, let $B_\alpha = \{x_\beta \mid \alpha \preceq \beta \}$ and show
        that $\{B_\alpha \}$ has the finite intersection property.

        To prove $\Leftarrow$, let $\mathcal A$ be a collection of closed sets having the finite
        intersection property, and let $\mathcal B$ be the collection of all finite intersections
        of elements of $\mathcal A$, partially ordered by reverse inclusion.}
\end{proof}

\end{document}
